Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.Rmd
clean up the paper and fill in the exp design

Heter Experiment design here

@@

# Old

## Overview of the Data

We collected 400 lineup evaluations made by 20 participants in experiment I and 880 lineup evaluations made by 44 participants in experiment II. In total, 442 unique lineups were evaluated by 64 subjects. In experiment I, one of the participants skipped all 20 lineups. Hence, the submission was rejected and removed from the dataset. In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment. Therefore, the submission was kept.

## Power comparision

1. power (visual test vs. conventional test)
(visual test most different one (everything test, any departure))
plot figure in a paper, desc, exp
2. investigate the difference (gap), give examples
3. conventional is too sensitive
4. make conventional less sensitive (vary alpha)

To model the power of visual test, 10 logistic regression were fit for different number of evaluations ranged from one to five and two different types of simulation setting. All 10 models used natural logarithm of the effect size as the only fixed effect, and whether the test successfully rejects the null hypothesis as the response variable. Given the way we define the effect size, it was expected that with larger effect size, both conventional test and visual test will have higher probability in rejecting the null hypothesis when it is not true. The modelling result summarized in \ref{tab:powerglmcubic} and \ref{tab:powerglmheter} aligned with the expectation as the coefficients of natural logarithm of the effect size are positive and significant across all 10 models.

Figure \ref{fig:power-com} illustrates the fitted models, while providing the local constant estimate of the power of F-test and Breusch–Pagan test for comparison. Data for the conventional test is simulated under the model setting described in section ... and 5000000 samples are drawn for both cubic and heteroskedasticity model. From Figure \ref{fig:power-com}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model.

For heteroskedasticity model, this phenomenon was more obvious as the power of visual tests with evaluations greater than two were always greater than those with evaluations smaller than two.

For cubic model, the separation between curves was small. The estimated power of visual tests with three to five evaluations were almost identical to each other in regards of effect size. In addition, all five curves peaked at one as effect size increased, suggesting that identification of non-linearity as a visual task can be completed reliably by human as long as the departure from null hypothesis is large enough.

As shown in Figure \ref{fig:power-com}, both F-test and Breusch–Pagan test generally possessed greater power than visual test. A visual tests is a collection of test against any alternatives that would create visual discoverable features, while a conventional test is usually targeting at a pre-specified alternative. Considering the data generating process of the model defect was known and controlled in this research, where all other alternatives have been eliminated except the one we concerned, the result was suggested that conventional tests were more sensitive to violations of linearity and homoscedasticity assumption than visual tests.

It was also found that there was a noticeable gap between curves of the conventional test and the visual test at around $log(\text{effect size}) = 0$ for the cubic model and $log(\text{effect size}) = 2.5$ for the heteroskedasticity model, where the differences in power were greater than 0.6. We further analysed the lineups with correspoding effect sizes. Figure \ref{fig:cubic-hard} and \ref{fig:heter-hard} showed that human was indeed hard to identify the patterns at this level of difficulty. The visual difference between the true data plot and null plots were almost unnoticeable.

<!-- sample 1, change values of a and b, plot the data  -->

## Effect of parameters on power of the visual test

The previous section focuses on the change of effect size relative to the power of the visual test. However, effect size is only a one dimensional summarisation of parameters used in data simulation. Individual factor embedded in the simulation process should also be analysed.

In cubic model, two major factors that influencing the strength of the signal are $a$ and $b$. Figure \ref{fig:power-com-cubic-a} and \ref{fig:power-com-cubic-a} illustrates 30 different logistic regressions fit for different number of evaluations and different number of observations $n$. The regressor used in these models was $|a|/\sigma$ since the noise level $\sigma$ needed to be taken into account. From the figures, we can observe ...

```{r power-vs-log-effect-size}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  ggplot() +
  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  xlab("Natural logarithm of effect size") +
  ylab("Power")
```

```{r power-vs-log-effect-size-given-x-dist}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  ggplot() +
  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  facet_wrap(~x_dist) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  ggtitle("Power comparison conditional on the distribution of X")
```

```{r power-of-visual-test-given-x-dist}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  ggplot() +
  geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  ggtitle("Power of visual test conditional on the distribution of X")
```

```{r power-vs-log-effect-size-given-shape}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  ggplot() +
  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  facet_wrap(~shape) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  ggtitle("Power comparison conditional on the shape of the Hermite polynomials")
```

```{r power-of-visual-test-given-shape}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>%
  ggplot() +
  geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  ggtitle("Power of visual test conditional on the shape of the Hermite polynomials")
```

```{r power-vs-log-effect-size-given-number-of-observations}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  ggplot() +
  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  facet_wrap(~n) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  ggtitle("Power comparison conditional on the number of observations")
```

```{r power-of-visual-test-given-number-of-observations}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), n = factor(n)) %>%
  ggplot() +
  geom_smooth(aes(log(effect_size), reject, col = n), method = "glm", method.args = list(family = binomial), se = FALSE) +
  xlab("Natural logarithm of effect size") +
  ylab("Power") +
  ggtitle("Power of visual test conditional on the number of observations")
```

```{r}
polynomials %>%
  group_by(lineup_id) %>%
  summarise(across(c(type:p_value), first)) %>%
  filter(lineup_id < 577) %>%
  mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>%
  ggplot() +
  geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.05)), aes(log(effect_size), reject, col = "Conventional test:0.05"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.01)), aes(log(effect_size), reject, col = "Conventional test:0.01"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.005)), aes(log(effect_size), reject, col = "Conventional test:0.005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0005)), aes(log(effect_size), reject, col = "Conventional test:0.0005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0000005)), aes(log(effect_size), reject, col = "Conventional test:0.0000005"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  xlab("Natural logarithm of effect size") +
  ylab("Power")
```

>>>>

Experiment II is designed to study the ability of human subjects to detect the appearance of a heteroskedasticity pattern under a simple linear regression model setting:

\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}

where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.

The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment:

\begin{equation} 
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

Model misspecification presents since the assumption about the constant error variance is violated.

Experiment data were simulated using three different shapes ($a$ = -1, 0, 1), five different values of $b$ (b = 0.25, 1, 4, 16, 64), three different sample sizes (n = 50, 100, 300) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.

The values of $a$ was chosen so that different shapes of heteroskedasticity were included in the residual plot. These include triangle shape, butterfly shape and inverse triangle shape.

```{r different-shape-of-heter}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

ori_dat <- heter_model(a = -1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300)

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote("Triangle shape")) + 
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("Butterfly shape")) + 
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("Inverse triangle shape")) + 
  theme_light() -> p3

patchwork::wrap_plots(p1, p2, p3, ncol = 3)
```

The range of $b$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.

@@

<!-- # Old -->

<!-- ## Overview of the Data -->

<!-- We collected 400 lineup evaluations made by 20 participants in experiment I and 880 lineup evaluations made by 44 participants in experiment II. In total, 442 unique lineups were evaluated by 64 subjects. In experiment I, one of the participants skipped all 20 lineups. Hence, the submission was rejected and removed from the dataset. In experiment II, there was a participant failed one of the two attention checks, but there was no further evidence of low-effort throughout the experiment. Therefore, the submission was kept. -->

<!-- ## Power comparision -->

<!-- 1. power (visual test vs. conventional test) -->
<!-- (visual test most different one (everything test, any departure)) -->
<!-- plot figure in a paper, desc, exp -->
<!-- 2. investigate the difference (gap), give examples -->
<!-- 3. conventional is too sensitive -->
<!-- 4. make conventional less sensitive (vary alpha) -->

<!-- To model the power of visual test, 10 logistic regression were fit for different number of evaluations ranged from one to five and two different types of simulation setting. All 10 models used natural logarithm of the effect size as the only fixed effect, and whether the test successfully rejects the null hypothesis as the response variable. Given the way we define the effect size, it was expected that with larger effect size, both conventional test and visual test will have higher probability in rejecting the null hypothesis when it is not true. The modelling result summarized in \ref{tab:powerglmcubic} and \ref{tab:powerglmheter} aligned with the expectation as the coefficients of natural logarithm of the effect size are positive and significant across all 10 models. -->

<!-- Figure \ref{fig:power-com} illustrates the fitted models, while providing the local constant estimate of the power of F-test and Breusch–Pagan test for comparison. Data for the conventional test is simulated under the model setting described in section ... and 5000000 samples are drawn for both cubic and heteroskedasticity model. From Figure \ref{fig:power-com}, it can be observed that the fitted power of visual test increased as the number of evaluations increased for both cubic and heteroskedasticity model. -->

<!-- For heteroskedasticity model, this phenomenon was more obvious as the power of visual tests with evaluations greater than two were always greater than those with evaluations smaller than two.  -->

<!-- For cubic model, the separation between curves was small. The estimated power of visual tests with three to five evaluations were almost identical to each other in regards of effect size. In addition, all five curves peaked at one as effect size increased, suggesting that identification of non-linearity as a visual task can be completed reliably by human as long as the departure from null hypothesis is large enough. -->

<!-- As shown in Figure \ref{fig:power-com}, both F-test and Breusch–Pagan test generally possessed greater power than visual test. A visual tests is a collection of test against any alternatives that would create visual discoverable features, while a conventional test is usually targeting at a pre-specified alternative. Considering the data generating process of the model defect was known and controlled in this research, where all other alternatives have been eliminated except the one we concerned, the result was suggested that conventional tests were more sensitive to violations of linearity and homoscedasticity assumption than visual tests.  -->

<!-- It was also found that there was a noticeable gap between curves of the conventional test and the visual test at around $log(\text{effect size}) = 0$ for the cubic model and $log(\text{effect size}) = 2.5$ for the heteroskedasticity model, where the differences in power were greater than 0.6. We further analysed the lineups with correspoding effect sizes. Figure \ref{fig:cubic-hard} and \ref{fig:heter-hard} showed that human was indeed hard to identify the patterns at this level of difficulty. The visual difference between the true data plot and null plots were almost unnoticeable.  -->

<!-- <!-- sample 1, change values of a and b, plot the data  --> -->

<!-- ## Effect of parameters on power of the visual test -->

<!-- The previous section focuses on the change of effect size relative to the power of the visual test. However, effect size is only a one dimensional summarisation of parameters used in data simulation. Individual factor embedded in the simulation process should also be analysed.  -->

<!-- In cubic model, two major factors that influencing the strength of the signal are $a$ and $b$. Figure \ref{fig:power-com-cubic-a} and \ref{fig:power-com-cubic-a} illustrates 30 different logistic regressions fit for different number of evaluations and different number of observations $n$. The regressor used in these models was $|a|/\sigma$ since the noise level $\sigma$ needed to be taken into account. From the figures, we can observe ... -->

<!-- ```{r power-vs-log-effect-size} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") -->
<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-x-dist} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   facet_wrap(~x_dist) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power comparison conditional on the distribution of X") -->
<!-- ``` -->

<!-- ```{r power-of-visual-test-given-x-dist} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power of visual test conditional on the distribution of X") -->
<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-shape} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   facet_wrap(~shape) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power comparison conditional on the shape of the Hermite polynomials") -->
<!-- ``` -->

<!-- ```{r power-of-visual-test-given-shape} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power of visual test conditional on the shape of the Hermite polynomials") -->
<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-number-of-observations} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   facet_wrap(~n) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power comparison conditional on the number of observations") -->
<!-- ``` -->

<!-- ```{r power-of-visual-test-given-number-of-observations} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), n = factor(n)) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = n), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power of visual test conditional on the number of observations") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.05)), aes(log(effect_size), reject, col = "Conventional test:0.05"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.01)), aes(log(effect_size), reject, col = "Conventional test:0.01"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.005)), aes(log(effect_size), reject, col = "Conventional test:0.005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0005)), aes(log(effect_size), reject, col = "Conventional test:0.0005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = mutate(conv_result, reject = as.numeric(p_value < 0.0000005)), aes(log(effect_size), reject, col = "Conventional test:0.0000005"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") -->
<!-- ``` -->

### 2022-08-31 paper_comparison/paper_comparison/paper_comparison.tex
clean up the paper and fill in the exp design

Heter Experiment design here

@@

\hypertarget{old}{%
\section{Old}\label{old}}

\hypertarget{overview-of-the-data}{%
\subsection{Overview of the Data}\label{overview-of-the-data}}

We collected 400 lineup evaluations made by 20 participants in
experiment I and 880 lineup evaluations made by 44 participants in
experiment II. In total, 442 unique lineups were evaluated by 64
subjects. In experiment I, one of the participants skipped all 20
lineups. Hence, the submission was rejected and removed from the
dataset. In experiment II, there was a participant failed one of the two
attention checks, but there was no further evidence of low-effort
throughout the experiment. Therefore, the submission was kept.

\hypertarget{power-comparision}{%
\subsection{Power comparision}\label{power-comparision}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  power (visual test vs.~conventional test) (visual test most different
  one (everything test, any departure)) plot figure in a paper, desc,
  exp
\item
  investigate the difference (gap), give examples
\item
  conventional is too sensitive
\item
  make conventional less sensitive (vary alpha)
\end{enumerate}

To model the power of visual test, 10 logistic regression were fit for
different number of evaluations ranged from one to five and two
different types of simulation setting. All 10 models used natural
logarithm of the effect size as the only fixed effect, and whether the
test successfully rejects the null hypothesis as the response variable.
Given the way we define the effect size, it was expected that with
larger effect size, both conventional test and visual test will have
higher probability in rejecting the null hypothesis when it is not true.
The modelling result summarized in \ref{tab:powerglmcubic} and
\ref{tab:powerglmheter} aligned with the expectation as the coefficients
of natural logarithm of the effect size are positive and significant
across all 10 models.

Figure \ref{fig:power-com} illustrates the fitted models, while
providing the local constant estimate of the power of F-test and
Breusch--Pagan test for comparison. Data for the conventional test is
simulated under the model setting described in section \ldots{} and
5000000 samples are drawn for both cubic and heteroskedasticity model.
From Figure \ref{fig:power-com}, it can be observed that the fitted
power of visual test increased as the number of evaluations increased
for both cubic and heteroskedasticity model.

For heteroskedasticity model, this phenomenon was more obvious as the
power of visual tests with evaluations greater than two were always
greater than those with evaluations smaller than two.

For cubic model, the separation between curves was small. The estimated
power of visual tests with three to five evaluations were almost
identical to each other in regards of effect size. In addition, all five
curves peaked at one as effect size increased, suggesting that
identification of non-linearity as a visual task can be completed
reliably by human as long as the departure from null hypothesis is large
enough.

As shown in Figure \ref{fig:power-com}, both F-test and Breusch--Pagan
test generally possessed greater power than visual test. A visual tests
is a collection of test against any alternatives that would create
visual discoverable features, while a conventional test is usually
targeting at a pre-specified alternative. Considering the data
generating process of the model defect was known and controlled in this
research, where all other alternatives have been eliminated except the
one we concerned, the result was suggested that conventional tests were
more sensitive to violations of linearity and homoscedasticity
assumption than visual tests.

It was also found that there was a noticeable gap between curves of the
conventional test and the visual test at around
\(log(\text{effect size}) = 0\) for the cubic model and
\(log(\text{effect size}) = 2.5\) for the heteroskedasticity model,
where the differences in power were greater than 0.6. We further
analysed the lineups with correspoding effect sizes. Figure
\ref{fig:cubic-hard} and \ref{fig:heter-hard} showed that human was
indeed hard to identify the patterns at this level of difficulty. The
visual difference between the true data plot and null plots were almost
unnoticeable.

\hypertarget{effect-of-parameters-on-power-of-the-visual-test}{%
\subsection{Effect of parameters on power of the visual
test}\label{effect-of-parameters-on-power-of-the-visual-test}}

The previous section focuses on the change of effect size relative to
the power of the visual test. However, effect size is only a one
dimensional summarisation of parameters used in data simulation.
Individual factor embedded in the simulation process should also be
analysed.

In cubic model, two major factors that influencing the strength of the
signal are \(a\) and \(b\). Figure \ref{fig:power-com-cubic-a} and
\ref{fig:power-com-cubic-a} illustrates 30 different logistic
regressions fit for different number of evaluations and different number
of observations \(n\). The regressor used in these models was
\(|a|/\sigma\) since the noise level \(\sigma\) needed to be taken into
account. From the figures, we can observe \ldots{}

\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-1.pdf}

\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-x-dist-1.pdf}

\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-x-dist-1.pdf}

\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-shape-1.pdf}

\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-shape-1.pdf}

\includegraphics{paper_comparison_files/figure-latex/power-vs-log-effect-size-given-number-of-observations-1.pdf}

\includegraphics{paper_comparison_files/figure-latex/power-of-visual-test-given-number-of-observations-1.pdf}

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-19-1.pdf}

>>>>

Experiment II is designed to study the ability of human subjects to
detect the appearance of a heteroskedasticity pattern under a simple
linear regression model setting:

\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}

where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vector of size \(n\).

The null regression model used to fit the realizations generated by the
above model is formulated exactly the same as the model used in the
first experiment:

\begin{equation} 
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

Model misspecification presents since the assumption about the constant
error variance is violated.

Experiment data were simulated using three different shapes (\(a\) = -1,
0, 1), five different values of \(b\) (b = 0.25, 1, 4, 16, 64), three
different sample sizes (n = 50, 100, 300) and four different
distribution of \(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
parameters used in this experiment is given in Table
\ref{tab:parameter-table}.

The values of \(a\) was chosen so that different shapes of
heteroskedasticity were included in the residual plot. These include
triangle shape, butterfly shape and inverse triangle shape.

\includegraphics{paper_comparison_files/figure-latex/different-shape-of-heter-1.pdf}

The range of \(b\), which is a factor controlling the strength of the
signal, was chosen so that different difficulty levels of lineups were
generated, and therefore, the estimated power curve would be smooth and
continuous.

@@

--\textgreater{}
